---
title: "Analysis of Teaching Methods and Reward Systems"
author: "JOSEPH MUTUA"
date: "`r Sys.Date()`"
output: pdf_document
---
#  Objective

The objective of this study is to determine whether patient skill test scores differ based on:

1. The **Teaching Method** used (Methods A, B, and C),
2. The **Reward System** applied (None, Praise, Tangible),
3. The **Interaction** between the teaching method and the reward system.

The experiment follows a Factorial Design with replication:
- **Factor 1 (Method):** 3 levels (A, B, C)
- **Factor 2 (Rewards):** 3 levels (None, Praise, Tangible)
- **Replications:** 4 scores per cell
- **Total Observations:** $3 \times 3 \times 4 = 36$



# Statistical Model

The model used for this analysis is:

$$Y_{ijk} = \mu + \alpha_i + \beta_j + (\alpha\beta)_{ij} + \epsilon_{ijk}$$

where:
- $\mu$ is the overall mean score.
- $\alpha_i$ is the effect of the $i^{th}$ Teaching Method.
- $\beta_j$ is the effect of the $j^{th}$ Reward System.
- $(\alpha\beta)_{ij}$ is the interaction effect between Method and Reward.
- $\epsilon_{ijk}$ is the random error term.


# Data Entry

```{r}

method <- rep(c("A", "B", "C"), each = 4, times = 3)
rewards <- rep(c("NONE", "PRAISE", "TANGIBLE"), each = 12)


scores <- c(
  
  52, 76, 60, 58, 
  58, 56, 68, 74,  
  58, 24, 32, 39,  
 
  60, 78, 75, 72,  
  60, 70, 74, 77,  
  56, 66, 54, 49,  
  
  98, 94, 96, 98,  
  78, 80, 84, 80,  
  72, 74, 76, 70   
)

medical_data <- data.frame(method, rewards, scores)
head(medical_data)

```


# Statistical Analysis (ANOVA)
We use the aov() function to test for main effects and the interaction.

```{r}
model <- aov(scores ~ method * rewards, data = medical_data)
summary(model)
```





# Results and Interpretation
Based on the ANOVA table generated from the data:

Teaching Method: 
$p = 2.27 \times 10^{-6}$ (Significant). Since the p-value is much smaller than $0.05$, we conclude that the choice of teaching method significantly impacts the patients' test scores.


Reward System: 

$p = 9.46 \times 10^{-9}$ (Significant). which is much smaller than $0.05$. , indicating that the type of reward (None, Praise, or Tangible) is a highly influential driver of performance.


Interaction (Method $\times$ Rewards):
$p = 0.101$ (Not Significant). Since $p > 0.05$, there is no evidence of a significant interaction. 


This suggests that the effectiveness of a teaching method does not depend on which reward system is being used; they work independently.


# Final Conclusion

At the 5% level of significance, we draw the following conclusions:

Teaching Methods:
We reject the null hypothesis ($H_0$) and conclude that there is a statistically significant difference in mean skill scores between the three teaching methods.

Reward Systems:
We reject the null hypothesis ($H_0$) and conclude that the type of reward significantly affects the patients' skill acquisition.

Interaction:
We fail to reject the null hypothesis ($H_0$). There is no significant interaction between the teaching method and the reward system.

# Summary:

Both the Teaching Method and the Reward System are critical factors in determining patient success. Because the interaction is not significant, practitioners can choose the best-performing teaching method and the most effective reward system independently to achieve the highest possible skill scores.